---
title: Working with empirical networks in socmod
author: Matt Turner
vignette: >
  %\VignetteIndexEntry{Working with empirical networks in socmod}
  %\VignetteEngine{quarto::quarto}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include: false
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r setup}
#| message: false
#| warning: false
library(socmod)
library(dplyr)
library(ggplot2)
```

Real-world datasets help ground diffusion models in empirical structure. `socmod`
ships with helpers for loading a friendship network collected by Coleman and Feld
(1982) and for simulating networks that reproduce the observed degree distribution.
This vignette shows how to

1. Inspect the network and quantify the friendship paradox
2. Visualize the structure with built-in plotting helpers
3. Seed an agent-based diffusion model directly from the empirical graph

## Load the Marketville friendship network

`socmod::get_feld_1991_network()` returns an undirected `igraph` object with 146
students. The dataset is bundled with the package so you can load it without
managing external files.

```{r}
marketville <- socmod::get_feld_1991_network()
marketville
```

We can verify basic properties—number of edges, degree distribution, and whether
any isolates are present.

```{r}
summary <- tibble::tibble(
  n_students = igraph::gorder(marketville),
  n_friendships = igraph::gsize(marketville),
  min_degree = min(igraph::degree(marketville)),
  median_degree = stats::median(igraph::degree(marketville)),
  max_degree = max(igraph::degree(marketville))
)
summary
```

## Quantify the friendship paradox

Many social networks exhibit the *friendship paradox*: most people have fewer
friends than their friends do. `compare_friendship_paradox()` computes how strong
this effect is for a given network.

```{r}
paradox <- socmod::compare_friendship_paradox(marketville)
paradox$summary
```

Use `plot_friendship_paradox()` to visualize each student's number of friends
against the average number of friends among their friends.

```{r}
socmod::plot_friendship_paradox(marketville)
```

## Visualize the network with `plot_network_adoption`

The same plotting utilities used for simulated models work on empirical graphs as
well. Here we colour nodes by whether they are initially assigned the adaptive
behavior.

```{r}
set.seed(123)
params_marketville <- socmod::make_model_parameters(
  graph = marketville,
  learning_strategy = socmod::contagion_learning_strategy,
  adaptive_fitness = 1.2,
  adoption_rate = 0.6,
  drop_rate = 0.1
)
model_marketville <- socmod::make_abm(params_marketville) %>%
  socmod::initialize_agents(initial_prevalence = 0.1, adaptive_fitness = 1.2)

plot_network_adoption(model_marketville, edgewidth = 0.4)
```

## Run a diffusion trial on the empirical network

Running simulations on real networks follows the same pattern as the toy examples.
After initialization we call `run_trial()` and then summarise the prevalence of the
adaptive behavior over time.

```{r}
trial_marketville <- socmod::run_trial(
  model_marketville,
  stop = socmod::fixated
)

socmod::summarise_outcomes(
  trial_marketville,
  outcome_measures = c("success_rate", "mean_fixation_steps")
)
```

```{r}
socmod::plot_prevalence(
  trial_marketville,
  tracked_behaviors = c("Adaptive", "Legacy")
)
```

## Simulate additional networks with the same degree profile

If you need more replicates with similar structure, call `simulate_feld_1991()` to
sample new graphs that match the Marketville degree distribution.

```{r}
set.seed(99)
simulated_network <- socmod::simulate_feld_1991()
igraph::degree(simulated_network) %>% head()
```

These tools make it straightforward to bridge empirical observations and model
dynamics within the `socmod` ecosystem.
